Advice on describing Bayesian analysis of neutron and X-ray reflectometry

### 2022-06-13 src/tex/ms.tex
Written up to likelihood section

\usepackage{showyourwork}
\usepackage{graphicx}
\usepackage[separate-uncertainty=true]{siunitx}
\sisetup{
    list-units = brackets,
    range-units = brackets,
    range-phrase = -,
    list-pair-separator = {, },
    list-separator = {, },
    list-final-separator = {, }
    }

@@

Reflectometry analysis can be described, in the most simplistic terms, as a comparison and refinement of a model based on some parameters, $\mathbf{x}$, to reproduce some reflectivity data set, $\mathbf{D}$. 
This refinement process involves comparing the model to the data and calculating some goodness-of-fit parameter or likelihood, $p(\mathbf{D} | \mathbf{x})$, and modifying the model to either minimise or maximise this, as appropriate. 
A commonly used goodness-of-fit parameter is the $\chi^2$ parameter \cite{nelson_refnx_2019} which is found as, 
%
\begin{equation}
    \chi^2 = \sum_{q=q_{\text{min}}}^{q_{\text{max}}}{\bigg[\frac{(R(q) - R(q)_{\text{m}})}{\sigma_R(q)}\bigg]^2}, 
\end{equation}
%
where, $R(q)$ is the experimental measured reflectivity, and $\sigma_R(q)$ is the associated uncertainty, at a given $q$ and $R(q)_{\text{m}}$ is the calculated reflectivity for the model at the same $q$.
The likelihood is related to the $\chi^2$ parameter, however, a larger likelihood is indicative of better agreement, 
%
\begin{equation}
    \ln[p(\mathbf{D} | \mathbf{x})] = -\frac{1}{2} \bigg(\chi^2 + \sum_{q=q_{\text{min}}}^{q_{\text{max}}}\ln{\big[2\pi\sigma_R(q)^2\big]}\bigg).
\end{equation}
%
The input for this refinement process is the model and some initial parameter values, which may be an absolute value or some parameter range, depending on the refinement algorithm. 
The output is a set of values for $\mathbf{x}$, potentially with associated error bars, where these are present they typically describe a standard deviation from the mean of a Gaussian probability distribution. 
This process implicitly assumes that the data is completely reduced, accounting for all experimental parameters, uncertainties are accurately described and the model can accurately describe the data.

The input required depends on a minimisation algorithm being used, with some algorithms requiring a single starting guess (such as traditional Newtonian methods) and others taking a range of potential values (more common in stochastic approaches, like differential evolution). 
The nature of these inputs define the results of the analysis, therefore it is of the utmost importance that these are shared as part of a publication describing the work. 
Furthermore, often the minimisation is performed with bounds in place, defining that the parameter values will lie within a given range. 
This range can be throught of as a having a prior probability distribution, where values of $x$ outside of this range have a probability of \num{0}. 
Even when a non-Bayesian approach is used in the analysis (i.e. Bayes theorem is not utilised), the result where bounds are set would be analogous to a Bayesian analysis with a uniform prior probability.

\section*{CRediT author statement}
\label{sec:credit}

@@

\begin{acknowledgments}
    We acknowledge.
\end{acknowledgments}

>>>>

\usepackage{showyourwork}
\usepackage{graphicx}
\usepackage{mhchem}
\usepackage[separate-uncertainty=true]{siunitx}
\sisetup{
    list-units = brackets,
    range-units = brackets,
    range-phrase = -,
    list-pair-separator = {, },
    list-separator = {, },
    list-final-separator = {, }
    }

@@

Reflectometry analysis can be described, in the most simplistic terms, as a comparison and refinement of a model based on some parameters, $\mathbf{x}$, to reproduce some reflectivity data set, $\mathbf{D}$. 
This refinement process involves comparing the model to the data and calculating some goodness-of-fit parameter or likelihood, $p(\mathbf{D} | \mathbf{x})$, and modifying the model to either minimise or maximise this, as appropriate. 
A commonly used goodness-of-fit parameter is the $\chi^2$ parameter which is found as \cite{nelson_refnx_2019}, 
%
\begin{equation}
    \chi^2 = \sum_{q=q_{\text{min}}}^{q_{\text{max}}}{\bigg[\frac{(R(q) - R(q)_{\text{m}})}{\sigma_R(q)}\bigg]^2}, 
\end{equation}
%
where, $R(q)$ is the experimental measured reflectivity, and $\sigma_R(q)$ is the associated uncertainty, at a given $q$ and $R(q)_{\text{m}}$ is the calculated reflectivity for the model at the same $q$.
The likelihood is related to the $\chi^2$ parameter, however, a larger likelihood is indicative of better agreement, 
%
\begin{equation}
    \ln[p(\mathbf{D} | \mathbf{x})] = -\frac{1}{2} \bigg(\chi^2 + \sum_{q=q_{\text{min}}}^{q_{\text{max}}}\ln{\big[2\pi\sigma_R(q)^2\big]}\bigg).
    \label{equ:likelihood}
\end{equation}
%
The input for this refinement process is the model and some initial parameter values, which may be an absolute value or some parameter range, depending on the refinement algorithm. 
The output is a set of values for $\mathbf{x}$, potentially with associated error bars, where these are present they typically describe a standard deviation from the mean of a Gaussian probability distribution. 
This process implicitly assumes that the data is completely reduced, accounting for all experimental parameters, uncertainties are accurately described and the model can accurately describe the data.

The input required depends on a minimisation algorithm being used, with some algorithms requiring a single starting guess (such as traditional Newtonian methods) and others taking a range of potential values (more common in stochastic approaches, like differential evolution). 
The nature of these inputs define the results of the analysis, therefore it is of the utmost importance that these are shared as part of a publication describing the work. 
Furthermore, often the minimisation is performed with bounds in place, defining that the parameter values will lie within a given range. 
This range can be throught of as a having a prior probability distribution, $p(\mathbf{x})$, where values of $x$ outside of this range have a probability of \num{0}. 
Even when a non-Bayesian approach is used in the analysis (i.e. Bayes theorem is not utilised), the result where bounds are set would be analogous to a Bayesian analysis with a uniform prior probability.

The optimised parameters from the minimisation algorithm, which depend on the particular algorithm, often include some statistical uncertainty. 
How these statistical uncertainties are found is beyond the scope of this work, but it is important to acknowledge that this uncertainty typically assumes that the probability distribution of the parameter is Gaussian in nature. 
This probability distribution is either the partial likelihood or posterior, the latter when some prior is included and Bayes theorem is applied, and describes our understanding of the parameter values. 
When Bayesian modelling is used and the prior is included, the posterior probability is found as, 
%
\begin{equation}
  p(\mathbf{x} | \mathbf{D}) \propto p(\mathbf{D} | \mathbf{x}) p(\mathbf{x}).
  \label{equ:bayes}
\end{equation}
%
Therefore, when Bayesian modelling is performed, the priors and likelihood are of fundamental important to the results that are obtained, the posterior, and therefore any scientific conclusions that are drawn.

The use of Bayesian inference can be valuable in the interpretation of reflectivity data, however, inconsistency in the description of the process will result in analysis that cannot be reproduced or easily understood. 
This can range from not giving the bounds of a uniform prior probability, to failing to share the complete sampling chain of a Markov chain Monte Carlo sampling or details of any autocorrelation analysis (the last of which, the authors of this work admit to being guilty of \cite{mccluskey_bayesian_2019}). 
In this letter, we outline best practice for reporting the results of Bayesian analysis for neutron and X-ray reflectometry, we hope that this work will engage others to carefully consider how they report this information. 
Furthermore, uptake of the approaches discussed herein will lead to greater clarity about the models and assumptions used in, and reproducibility of, our analyses.

\section{Prior}
\label{sec:prior}

The most common prior probability that is used for parameter is uniform between two values.
The used of a bounded parameter along with some tradiational $\chi^2$-minimisation method and a parameter with a uniform prior and a Bayesian maximum \emph{a-posteriori} approach will lead to the same result. 
For priors that are uniform is it important that the upper and lower bounds are reported, this can be achieved with a simple table to be included in the article or supplementary information, we show an example of this in Table~\ref{tab:bounds}. 
%
\begin{table}
    \caption{An example of the presenting uniform priors in a tabular format. Reproduced from \cite{mccluskey_general_2020}, where each parameter was either constrained to a given value or sampled within the prior range.}
    \label{tab:bounds}
    \begin{tabular}{c| c | c}
        Parameter & Constrained Value & Prior Range \\ 
        \hline
        $d_h$/\si{\angstrom} & $10.0$ & $[8.0, 16.0)$\\
        $V_h$/\si{\angstrom^3} & $339.5$ & $[300.0, 380.0)$\\
        $d_t$/\si{\angstrom} & $21.0$ & $[10.0, 26.0)$\\
        $\phi_t$ & $1.0$ & $[0.5, 1.0)$\\
        $V_t$/\si{\angstrom^3} & $850.4$ & $[800.0, 1000.0)$\\
        $\sigma$/\si{\angstrom} & $2.9$ & $[2.9, 5.0)$\\
    \end{tabular}
\end{table}
%

Currently, the use of non-uniform, informative priors is less common in the analysis of reflectometry. 
However, the increasing popularity of Bayesian methods and interest in using complementary methods for analysis means that these are likely to become more popular in the coming years. 
Here we will define two potential types of informative prior probabilities, those that can be described with some mathematical function and those that cannot, for example arising from the application of a sampling-based analysis of a complementary technique.

When a prior probability can be described with a mathematical function, this should be done by providing this function in the clearest possible language. 
For example, if the prior is taken from a single complementary measurement that is deinfed as a value with some uncertainty, which represents a normal distribution with a mean and standard deviation, this information can be provided. 
This is shown in Figure~\ref{fig:prior} for the density of silicon nitride (\ce{Si3N4}) produced by atomic layer deposition \cite{knoops_atomic_2015}) which is used to inform the value for a scattering length density for some layer of the material. 
Such a prior probability distribution could be described graphically as in Figure~\ref{fig:prior}, in prose as being ``normally distributed with a mean of \SI{2.9}{\gram\per\cubic\centi\meter} and a standard deviation of \SI{0.1}{\gram\per\cubic\centi\meter}'', mathematically as, 
%
\begin{equation}
  p(\rho_m) = \frac{1}{\sigma\sqrt{2\pi}}\exp\Bigg[-\frac{1}{2}\bigg(\frac{\rho_m-\mu}{\sigma}\bigg)^2\Bigg]
\end{equation}
%
where, $\mu=\SI{2.9}{\gram\per\cubic\centi\meter}$ and $\sigma=\SI{0.1}{\gram\per\cubic\centi\meter}$, or more concisely $p(\rho_m) \sim \mathcal{N}(\mu=\SI{2.9}{\gram\per\cubic\centi\meter}, \sigma=\SI{0.1}{\gram\per\cubic\centi\meter})$. 
The same descriptive approach could be taken for any common statistical distribution, including log-normal or truncated normal distributions. 
%
\begin{figure}
  \script{prior.py}
  \includegraphics[width=\columnwidth]{figures/prior.pdf}
  \caption{
    A prior probability distribution for \ce{Si3N4} with a density of $\rho_m=\SI{2.9\pm0.1}{\gram\per\cubic\centi\meter}$. 
  }
  \label{fig:prior}
\end{figure}
%

In the case that the prior distribution cannot be described easily with some mathematical function, for example, if it is a multimodal model result from some other sampling method, then the chain from the prior sampling should be given. 
The chain is all of the samples investigated in the sampling should be share, although in the case that this chain is very large a subsampled object shared, in which case, the autocorrelation analysis performed should be described (see Section~\ref{sec:posterior} for complete discussion of this).
In order to use such a prior probability in Bayesian analysis, it is necessary that some functional description of the prior is generated, most commonly this will be some kernel density estimation, when this is used it is also necessary to state the structure of the kernel being used. 
An example of this is shown in Figure~\ref{fig:samples}, where the prior probability for the volume of a phospholipid tail could be found from molecular dynamics simulation where there are three common conformers that the lipid is likely to have.
%
\begin{figure}
  \script{samples.py}
  \includegraphics[width=\columnwidth]{figures/samples.pdf}
  \caption{
    A prior probability distribution for a DPPC lipid that could arise from a molecular dynamics simulation (orange histrogram) and a Gaussian kernel density estimation for the probability distribution using a badwidth factor of \num{0.05} (blue line). 
  }
  \label{fig:samples}
\end{figure}
%

\section{Likelihood}
\label{sec:likelihood}

Bayes theorem (given in Equation~\ref{equ:bayes}) consists of the product of the prior and the likelihood. 
The former describes our current understanding of the system before we conduct some experiment, while the latter describes the understanding that we gain from the data. 
Therefore, it is necessary that we state how the likelihood for a given model and data is calculated, in particular because although Equation~\ref{equ:likelihood} is a common approach, it assumes a normally distributed uncertainty. 
While a normally distributed uncertainty is the most common, it may not be accurate in all circumstances, for example, when low numbers of counts are present, a Poisson uncertainty may offer a more accurate description.

\section{Posterior}
\label{sec:posterior}

\section{Conclusion}
\label{sec:conclusions}

ZENODO.

\section*{CR\lowercase{e}d\lowercase{i}T author statement}
\label{sec:credit}

@@

\begin{acknowledgments}
    All authors gratefully acknowledgment the contribution of Rev. T. Bayes (1701-1761). 
\end{acknowledgments}
